%% Language-Based Security project proposal template
%% 2022-02-07
%% To use this template:
%% - Remove \instruction commands and fill-in with your text
%% - Include your names in the author command
%%
%% - Compile with Makefile
%% - Edit literature.bib to include relevant literature.
%%
%% - Submit _only_ the resulting PDF file.
%%



\documentclass[10pt]{article}
\input{prelude}
\input{instructions}

\title{
   Assignment 1
}

\groupNumber{18}
\author{Firas Saleh \and Henrique Luz \and Marina Toledo }

\begin{document}
\maketitle
\thispagestyle{empty}

\section{Report}
\label{sec:descrpition}

\instruction{
Describe your solution, addressing the following questions
\begin{itemize}
  \item How did you approach the assignment?
  \item What did you try?
  \item What worked and what did not work?
  \item What have you discovered when you decrypted the file?
  \item What is the security model in this assignment? What is the threat model? What security requirements are being violated? 
  \item What have you learned from this exercise?
  \item Any other insights?
\end{itemize}
}

\section{Introduction}

SQL injection remains one of the most prevalent and dangerous vulnerabilities in web applications.
This assignment investigates a blind SQL injection attack against a password reset service
implemented in Node.js with Express and sqlite3. Unlike traditional SQL injection attacks where
query results are directly visible, blind SQL injection requires attackers to infer information
through observable side effects, such as application behavior or response timing.


The target application accepts a username and queries the database using string concatenation
without parameterization, creating an SQL injection vector. The service always returns the same
message regardless of query success, eliminating content-based information disclosure. However,
by leveraging timing differences induced by database operations, we can extract sensitive data
character by character.

This report details our methodology, optimizations, and theoretical analysis
of this timing-based attack.

\section{Vulnerability Analysis}

\subsection{Attack Surface}

The vulnerable endpoint constructs SQL queries through direct string concatenation:
\[
  \texttt{"select * from users where username='" + user + "'"}
\]
This construction allows arbitrary SQL code injection when the \texttt{user} variable contains malicious
input. The absence of input validation, parameterized queries, or prepared statements creates a
critical security vulnerability.

\subsection{Database Schema}
The database contains a \texttt{users} table with three columns: \texttt{username}, \texttt{passwordhash}, and
\texttt{key}. The \texttt{key} column stores the user's private PGP decryption key in ASCII format.

\subsection{Attack Objective}
The goal is to retrieve the private key stored in the \texttt{key} field for user \texttt{admin}. The key follows the
standard PGP private key block format, beginning with \texttt{-----BEGIN PGP PRIVATE KEY BLOCK-----} and ending with \texttt{-----END PGP PRIVATE KEY BLOCK-----}. Since the
application provides no direct output channel for query results, we must employ an inference-based approach.

\section{Methodology}

\subsection{Timing-Based Boolean Inference}
Our approach is inspired by Meer and Slaviero's work on timing-based attacks. The fundamental
technique involves constructing SQL queries that conditionally introduce time delays based on
boolean predicates. By measuring response times, we can determine whether specific conditions
are true or false.
We inject the following payload into the username field:
\[
  \texttt{admin' AND condition AND sleep(sleeptime) --}
\]
Our custom sleep function creates a SQL condition that forces the database to generate 
a large random blob. The function randomblob(n) in SQLite3 generates random data of 
size n bytes. By requesting a large blob (with size controlled by the sleeptime 
parameter), we force the database to perform significant computational work.
We calibrated the blob size to produce delays of approximately 
sleeptime seconds. The sleep condition compares this blob to a ranodm value 
(e.g., randomblob(sleeptime*100000000) = '1234'), which evaluates to false, but 
the computational cost lies in the blob generation itself, not the comparison.

Due to short-circuit evaluation of the AND operator in our injection payload 
(condition AND sleep(sleeptime)), if condition evaluates to false, the database 
returns immediately without generating the blob. If condition is true, the database 
must generate the blob before evaluating the comparison, introducing a measurable delay.

\subsection{Initial Approach: Brute Force Character Matching}

Our initial approach involved testing each character position against all possible ASCII characters
(0-127). For each position i, we would inject:
\[
  \texttt{substr(key, i, 1) = 'c'}
\]
for each candidate character c. With uniform distribution of characters, this requires an average of
64 requests per character (128/2 expected comparisons until finding the match).
Assuming the target key has length $n$ characters, and denoting the baseline response time as $t_{base}$
and the delay time as $t_{delay}$, the expected total time for brute force is:
\[
  T_{brute} = n \times (63 \cdot t_{base} + t_{delay})
\]

\subsection{Optimization: Binary Search}
To significantly reduce the number of requests, we implemented a binary search algorithm. Instead
of testing each character value, we test whether the character's Unicode value is less than or equal
to a midpoint:
\[
\texttt{unicode(substr(key, i, 1)) <= mid}
\]
The algorithm maintains bounds $[low, high]$ initialized to $[0, 127]$ for ASCII characters.
At each iteration, we test the midpoint $mid = (low + high) / 2$. If the condition is true
(indicated by a long response), we update $high = mid$; otherwise, $low = mid + 1$. The
search terminates when $low == high$.

This reduces the worst-case number of queries per character to
\[
\lceil \log_2(128) \rceil = 7,
\]
with an average of approximately
\[
\log_2(128) - 1 = 6
\]
queries. The expected time becomes:
\begin{quote}
  $T_{\text{binary}} = L \cdot 6 \cdot t_{\text{base}} \quad \text{(assuming ideal conditions)}$
\end{quote}

However, this assumes we never encounter delayed responses. In practice, we must account for
true positive responses and verification overhead.

\subsection{Challenges and Solutions}

\subsubsection{Network Variability}
Network latency and server load create non-deterministic response times. Even when the
condition is false, response time may occasionally exceed our threshold $t_{\text{threshold}}$, causing false
positives. If we denote the probability of such spurious delays as $p$, then without mitigation, our
results would be unreliable.

\subsubsection{Initial Solution: Retry on True}
Our first mitigation strategy involved retrying the same condition $n$ times whenever we detect a
delayed response. If all $n$ retries also show delays, we conclude the condition is truly true. This
reduces the false positive probability to $p^n$.

However, this approach severely degrades performance. In binary search, we expect
approximately 3 true responses per character on average (half of the 6 queries return true). Each
true response now requires $n + 1$ delayed responses (the initial detection plus $n$ retries), giving
expected time per character:
\begin{quote}
  $T_{\text{char}} = 3 \cdot t_{\text{base}} + 3 \cdot (n+1) \cdot t_{\text{delay}}$
\end{quote}

With $n = 3$ retries and $t_{\text{delay}} = 2$ seconds, this becomes approximately 24 seconds per character,
making the optimization counterproductive.

\subsubsection{Improved Solution: Negative Condition Validation}
We developed a more efficient validation strategy: when detecting a potential true response,
immediately test the negation of the condition. We inject:

\begin{quote}
\texttt{admin' AND NOT(condition) AND randomblob(2*sleeptime*100000000)='1' --}
\end{quote}

If the original condition is truly true, the negation must be false, resulting in a fast response. If we
observe a fast response to the negation, we confirm the original was genuinely true. If the negation
also produces a delay, the original detection was a false positive caused by network issues.

Note that we double the sleep time in the negation test ($2 \cdot \text{sleeptime}$). This accounts for
potentially persistent network congestion—if delays are due to network issues rather than query
logic, the doubled sleep ensures we still observe a timeout, preventing false confirmations.

\subsubsection{Adaptive Sleep Time}
Network conditions change dynamically. We implemented adaptive threshold adjustment: when
detecting false positives (negation test fails), we double the \texttt{sleeptime} parameter. When
conditions stabilize (several consecutive tests without false positives), we gradually reduce
\texttt{sleeptime} to maintain efficiency. This balances robustness against varying network conditions
with minimal performance overhead.


%% References
%% - Edit literature.bib to add references
%%
\bibliographystyle{abbrvnat}
\bibliography{literature}


\end{document}
